Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-02-01 paper/zhang-cook-laa-langrene-menendez.tex
finish proofreading

{\centering \includegraphics[width=0.5\linewidth,height=0.25\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}

@@

\noindent where \(I_d\) is the \(d\)-dimensional identity matrix and the
constraint requires the projection bases \(\mathbf{A}\) to be
orhtonormal.

Several features of this optimisation are worth-noticing. 1) The
optimisation is constrained and the space of the bases is a
d-dimensional sphere. 2) The objective function may not be
differentiable for a constructed index function. 3) Although finding the
global maximum is the goal of an optimisation problem, it is also
interesting to inspect local maximum in projection pursuit since it
could present unexpected interesting projections. 4) The optimisation
procedure needs to be easy to compute since the tour animation needs to
be played in real-time during the optimisation.

@@

Three main elements are recorded for the projection pursuit optimisers:
1) projection bases: \(\mathbf{A}\), 2) index values: \(I\), and 3)
State: \(S\), which labels the observation with detailed stage in the
optimisation. For optimiser SA and SAJO, possible values of the state
include \texttt{random\_search}, \texttt{new\_basis}, and
\texttt{interpolation}. pseudo derivative search has a wider variety of
state including \texttt{new\_basis}, \texttt{direction\_search},
\texttt{best\_direction\_search}, \texttt{best\_line\_search}, and
\texttt{interpolation}.

Multiple iterators are also needed to index the data collected at
different levels: \(t\) being a unique identifier that prescribes the
natural ordering of each observation while \(j\) and \(l\) being the
counter of the outer and inner loop, respectively, in Algorithm
\ref{random-search}, \ref{simulated-annealing} and \ref{search-geodesic}
above. Other parameters of interest recorded include
\(V_1 = \text{method}\) that tags the name of the optimiser, and
\(V_2 = \text{alpha}\) that indicates the sampling neighbourhood size. A
matrix notation of the data structure is presented in Equation
\ref{eq:data-structure}.

@@

\noindent where \(T^{\prime} = T + k_{J}+ l_{J+1}\). Note that there is
no output in iteration \(J + 1\) since the optimiser cannot find a
better basis in the last iteration and the algorithm terminates. The
final basis found is \(A_T\) with the highest index value \(I_T\).

@@

The construction of diagnostic plots adopts the core concept in ggplot2:
grammar of graphics \citep{wickham2010layered}. In grammar of graphics,
plots are not produced via calling the commands, named by the appearance
of the plot, i.e., boxplot and histogram, but via the concept of stacked
layers. Seeing plots as stacked layers empowers us to composite
diagnostic plots with an emphasis on any variable in the data object
without the redundancy of creating different commands for the same type
of plots that highlights on different variables.

@@

A starting point of diagnosing an optimiser is to understand how many
search the optimiser has conducted. One may want to simply plot the
index value of the search points across its natural order. A point
geometry may work well if each iteration has a similar number of points,
however, the plot will be disproportionate if some iterations have
considerably more points than other - it over-emphasizes the iterations
that have more search points! An alternative is to summarise the search
in each iteration using boxplot where each iteration will be spaced out
equally. Occasionally, one may still want to switch back to point
geometry if the number of points is small and this can be achieved via
the \texttt{cutoff} argument. Additional annotation can be added with
new layers, for example, the number of points searched in each iteration
can be added as text label at the bottom of each iteration; the anchor
bases to interpolate are connected and highlighted in a larger size; and
the colour of the last iteration is in a grey scale to indicate no
better basis found in this iteration.

Figure \ref{fig:toy-search} shows the searches of two different
optimisers: SA (left) and SAJO (right). Both optimisers quickly find
better bases at the first few iterations, then take longer in the later
iterations, and finally finish when the maximum number of evaluation is
reached. The anchor bases, the ones with the highest index value in each
iteration, always have an increased index value in the optimiser SA
while this is not the case for SAJO. This feature gives SA an advantage
in simple example, like the one presented while the jump out component
in SAJO allows for a more holistic search in more complicated examples.

@@

\caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes indexwith two optimisers: SA and PD. All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}\label{fig:toy-pca}
\end{figure}
\end{Schunk}

Apart from checking the progression of an optimiser, looking at how the
points evaluated by the optimisers look like in the space is another
interest. Given the orthonormality constraint, the space of projection
bases \(\mathbf{A}_{p \times d}\) is a \(p \times d\) dimension sphere
and dimension reduction methods, i.e.~principal component analysis is
applied first to project all the bases onto a 2D space. In a projection
pursuit guided tour optimisation, there are various type of bases
involved: 1) The start basis; 2) The anchor bases that have the highest
index value in each iteration; 3) The search bases that an optimiser
evaluated to produce the anchor bases; 4) The interpolating bases on the
interpolation path; and finally 5) the end basis. The importance of
these basis differs with the most important ones being the start,
interpolating and end bases. Anchor and search bases can be turned on
with argument \texttt{details\ =\ TRUE}. Sometimes, two optimisers can
start with the same basis but finish with bases of opposite sign. This
happens because the projection is invariant to the sign difference of
the bases and so does the index value, however, this creates
difficulties for comparing end bases. A flip sign device is implemented
to flip the sign of all the bases in one optimisation if the situation
described above happens and to ensure different routines finish close to
each other.

Several annotations have been made to help understanding this plot. The
original space of the bases is a high dimensional sphere and random
bases on the sphere can be generated via the CRAN package
\CRANpkg{geozoo} \citep{geozoo}. Along with the bases recorded during
the optimisation and a zero basis, PCA is performed to get the first two
PC coordinates of all the bases. The search space in the 2D space is a
circle with the origin being the PC coordinates of the zero matrix, and
radius estimated as the largest distance between the origin and all the
bases. The theoretical best basis is known with simulated data and can
be labelled for comparing the how close the ending basis to the
theoretical one for different optimisers. Various aesthetics, i.e.~size,
alpha and colour, are applicable to emphasize the crucial elements and
adjust for the presentation. For example, anchor points and search
points are less important and a smaller size and alpha is used; The
interpolation paths are more informative if alpha increases from start
to finish (that is, the path is more opaque towards the end).

Figure \ref{fig:toy-pca} presents the PCA plot of a 1D projection of a
5-variable dataset, \code{boa5}, using the holes index with two
optimisers, PD and SA, being compared. Both optimisers get close to the
theoretical bases, annotated by "*" and in this particular example
pseudo derivative search gets closer. One can also appreciate the nature
of the two different searches: PD first evaluates points in a small
neighbourhood for a promising direction, while SA evaluates points
randomly in the search space to search for the next target. There are
dashed lines annotated for SA and it describes the interruption of the
interpolation, which has been briefly mentioned in Figure
\ref{fig:toy-interp} and will be discussed in details in Section
\ref{monotonic}.

@@

Animation is another display to show how the search progresses from
start to finish in the space and an \texttt{animate\ =\ TRUE} argument
is used to enable the animation. Figure \ref{fig:toy-pca-animated} shows
six frames from the animation of the PCA plot in Figure
\ref{fig:toy-pca}. An additional piece of information one can learn from
this animation is that the SA search finds its end basis quicker than
the PD search since SA has finished in the 5th frame while PD is still
making more progression.

@@

As mentioned in Section \ref{toy-pca}, the original \(p \times d\)
dimension space can be simulated via randomly generated bases via
\CRANpkg{geozoo} \citep{geozoo} package. A tour plot rotates these
random bases and the ones collected by the optimisation in the original
space and hence provides a stereoscopic view of the search. While the
PCA plot projects the bases from the direction that maximises the
variance, the tour plot display of the original high dimensional space
from various directions using animation. Figure \ref{fig:toy-tour} shows
the tour plot of the same two optimisations in the 5D space.

@@

Random variables with different distributions have been simulated and
the distributional form of each variable is presented in Equations
\ref{eq:sim-norm} to \ref{eq:sim-x7}. Variable \texttt{x1}, \texttt{x8},
\texttt{x9} and \texttt{x10} are normal noise with zero mean and unit
variance, and \texttt{x2} to \texttt{x7} are normal mixtures with varied
weights and locations. All the variables have been scaled to have an
overall unit variance before projection pursuit.

@@

Once the final basis has been found by an optimiser, one may want to
push further in the close neighbourhood to see if an even better basis
can be found. A polish search takes the final basis of an optimiser as
the start of a new guided tour to search for local breakthrough. The
polish algorithm is similar to the SA with three main distinctions: 1)
100 rather than one candidate bases are generated each time in the inner
loop; 2) the search neighbourhood, \texttt{alpha}, is reduced in the
inner loop, rather than in the outer loop, once a set of candidate bases
fails to make an improvement; and 3) three more termination conditions
have been added to ensure the new basis generated is distinguishable
from the current in terms of the distance in the space, percentage
change in the index value, and neighbourhood:

@@

The index function before this section are all smooth, while this is not
the case for all. A 1D projection function based on the Kolmogorov test,
\(I^{nk}(n)\), compares the projected data, \(\mathbf{Y}_{n \times 1}\),
to a randomly generated normal distribution,
\(\mathcal{N}_{n \times 1}\), based on the empirical cumulated
distribution function (ECDF). Let \(F_{.}(n)\) be the ECDF function,
with two possible subscripts \(Y\) and \(\mathcal{N}\) representing the
projected and randomly generated data, and \(n\) denoting the number of
observation, the Normal Kolmogorov, \(I^{nk}(n)\), index is defined as:

\[I^{nk}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]\] With a
non-smooth index function, two research questions are raised to
understand

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  if any optimiser fails to optimise a non-smooth index; and
\item
  whether the optimisers can find the global optimum despite the
  presence of a local optimum
\end{enumerate}

@@

Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of
all three optimisers. As expected, none of the interpolated path is
smooth. There is barely any improvement made by PD, indicating its
failure in optimising non-smooth index. While SA and SAJO have managed
to get close to the index value of the theoretical best, trace plot
shows that it takes SAJO much longer to interpolate towards the final
basis. This long interpolation path is partially due to the fluctuation
in the early iterations, where the SAJO tends to generously accept
inferior basis before concentrating near the optimum. The PCA plot shows
the interpolation path and search points excluding the last termination
iteration. {[}A sentence for PD{]}. Comparing the amount of random
search done by SA and SAJO, it is surprising that although SAJO doesn't
actually carry that many of search as SA. Combine the insight from both
the trace and PCA plot, one can learn the two different search
strategies by SA and SAJO: SA frequently samples in the space and only
make a move when an improvement will be made, while SAJO first broadly
accepts basis in the space and only starts the extensive search in a
narrowed subspace. The specific decision of which optimiser to use will
depends on the index curve in the basis space but if the basis space is
highly non-smooth and spike, accepting inferior bases at first, as SAJO,
can lead to a more efficient search, in terms of the overall number of
points evaluated.

@@

The next experiment compares the performance of SA and SAJO when a local
maximum exists. Two search neighbourhood sizes: 0.5 and 0.7, are
compared to understand how a large search neighbourhood would affect the
final basis and the length of the search. Figure \ref{fig:kol-result}
shows 80 paths simulated using 20 seeds in the PCA plot, faceted by
optimiser and search size. With SA and a search size of 0.5, despite
being the simplest and fastest, SA fails to optimise in three instances
where they finish neither near the local nor the global optimum. With a
larger search size of 0.7, none of the seed fail and more seeds find the
global maximum. Comparing SA and SAJO for a search size of 0.5, SAJO
doesn't seem to improve the final basis found, despite having longer
interpolation paths. However, the denser path near the local maximum is
an indicator that SAJO is working hard to examine if there is any other
optimum in the space but the relative small search size has diminished
its ability to reach out to the true global maximum. With a larger
search size, almost all the seeds (16 out of 20) have found the global
maximum and some final bases are much closer to the theoretical best, as
compared to the three other cases. This indicates that SAJO, with a
reasonable large search size, is able to overcome the local optimum and
optimise close towards the underlying true optimum.

@@

The implementation of this projection has been divided into two
packages: the data collection object is implemented in the existing CRAN
package \CRANpkg{tourr} \citep{tourr} while the optimiser diagnostics
have been implemented in a new package, \pkg{ferrn}. When a guided tour
is run, the users can choose if the data from optimisation should be
collected via the \texttt{verbose} argument. Once the data object has
been obtained, the package, \pkg{ferrn}, can provide four diagnostic
plots as shown in Section \ref{vis-diag}. The structure of package
functionality has been listed below.

@@

  \begin{itemize}
  \tightlist
  \item
    \code{explore\_trace\_search()} produces summary plots in Figure
    \ref{fig:toy-search}
  \item
    \code{explore\_trace\_interp()}produces trace plots for the
    interpolation points in Figure \ref{fig:toy-interp}
  \item
    \code{explore\_space\_pca()} produces the PCA plot of projection
    bases on the reduced space in Figure \ref{fig:toy-pca}. Animated
    version in Figure \ref{fig:toy-pca-animated} can be turned on via
    the argument \texttt{animate\ =\ TRUE}
  \item
    \code{explore\_space\_tour()} produces animated tour view on the
    full space of the projection bases in Figure \ref{fig:toy-tour}.
  \end{itemize}
\item
  \code{get\_*()} extracts and manipulates certain components from the
  existing data object.

@@

  \begin{itemize}
  \tightlist
  \item
    \code{bind\_theoretical()} binds the theoretical best observation in
    simulated experiment
  \item
    \code{bind\_random()} binds randomly generated bases in the
    projection bases space to the data object
  \item
    \code{bind\_random\_matrix()} binds randomly generated bases and
    outputs in a matrix format
  \end{itemize}
\item
  Utilities

  \begin{itemize}
  \tightlist
  \item
    \code{add\_*()} are internal wrapper functions that facilitate the
    composition of PCA plot
  \item
    \code{theme\_fern()} and \code{format\_label()} for better display
    of the grid lines and axis formatting
  \item
    \code{clean\_method()} for clean up the name of the optimisers
  \item
    \code{botanical\_palettes} is a collection of colour palettes from
    Australian native plants Quantitative palettes include daisy,
    banksia and cherry and sequential palettes contain fern and acacia
  \item
    \code{botanical\_pal()} as the colour interpolator
  \item
    \code{scale\_color\_*()} and \code{scale\_fill\_*()} for scaling the
    colour and fill of the plot
  \end{itemize}
\end{itemize}

>>>>

{\centering \includegraphics[width=0.5\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/PhD/research/paper-tour-vis/img/tour_path}

@@

\noindent where \(I_d\) is the \(d\)-dimensional identity matrix and the
constraint requires the projection bases \(\mathbf{A}\) to be
orhtonormal. Several features of this optimisation are worth-noticing.
1) The optimisation is constrained and the space of the bases is a
d-dimensional sphere. 2) The objective function may not be
differentiable for a constructed index function. 3) Although finding the
global maximum is the goal of an optimisation problem, it is also
interesting to inspect local maximum in projection pursuit since it
could present unexpected interesting projections. 4) The optimisation
procedure needs to be easy to compute since the tour animation needs to
be played in real-time during the optimisation.

@@

Three main pieces of information are recorded for the projection pursuit
optimisers: 1) projection bases \(\mathbf{A}\), 2) index values \(I\),
and 3) state \(S\). For SA and SAJO, possible state include
\texttt{random\_search}, \texttt{new\_basis}, and
\texttt{interpolation}. Pseudo derivative (PD) has a wider variety of
state including \texttt{new\_basis}, \texttt{direction\_search},
\texttt{best\_direction\_search}, \texttt{best\_line\_search}, and
\texttt{interpolation}. Multiple iterators index the information
collected at different levels: \(t\) is a unique identifier prescribing
the natural ordering of each observation; \(j\) and \(l\) are the
counter of the outer and inner loop respectively. Other parameters of
interest recorded include \texttt{method} that tags the name of the
optimiser, and \texttt{alpha} that indicates the sampling neighbourhood
size for searching observations. A matrix notation of the data structure
is presented in Equation \ref{eq:data-structure}.

@@

\noindent where \(T^{\prime} = T + k_{J}+ l_{J+1}\). Note that there is
no output in iteration \(J + 1\) since the optimiser doesn't find a
better basis in the last iteration and terminates. The final basis found
is \(A_T\) with index value \(I_T\).

@@

The construction of diagnostic plots adopts the core concept of grammar
of graphics \citep{wickham2010layered} in ggplot2. In grammar of
graphics, plots are not produced by calling the commands, named by the
appearance of the plot, i.e., boxplot and histogram, but by the concept
of stacked layers. Seeing plots as stacked layers gives analysts the
freedom to composite plots with multiple elements on hands.

@@

A starting point of diagnosing an optimiser is to understand how many
search an optimiser has conducted. One may want to simply plot the index
value of the search points across its natural order, but a point
geometry may work well if each iteration has a similar number of points.
When some iterations have considerably more points than others, using a
point geometry over-emphasizes the iterations that have more search
points since these iterations will occupy the mast majority of the plot
space. An alternative is to summarise the search in each iteration using
boxplot and each iteration will then be spaced out equally.
Occasionally, one may still want to switch back to a point geometry if
the number of points is small in a particular iteration and this can be
achieved via the \texttt{cutoff} argument in the search plot function.
Additional annotations are added to facilitate better reading of the
plot and these includes\\
1) the number of points searched in each iteration can be added as text
label at the bottom of each iteration; 2) the anchor bases to
interpolate are connected and highlighted in a larger size; and 3) the
colour of the last iteration is in a grey scale to indicate no better
basis found in this iteration.

Figure \ref{fig:toy-search} shows an example of the search plot for SA
(left) and SAJO (right). Both optimisers quickly find better bases in
the first few iterations and then take longer to find one in the later
iterations. The anchor bases, the ones found with the highest index
value in each iteration, always have an increased index value in the
optimiser SA while this is not the case for SAJO. This feature gives SA
an advantage in this simple example to quickly find the optimum.

@@

\caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers: SA and PD. All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}\label{fig:toy-pca}
\end{figure}
\end{Schunk}

Apart from checking the search and progression of an optimiser, looking
at where the bases are positioned in the basis space is another
interest. Given the orthonormality constraint, the space of projection
bases \(\mathbf{A}_{p \times d}\) is a \(p \times d\) dimension sphere
and dimension reduction methods, i.e.~principal component analysis is
applied to first project all the bases onto a 2D space. In a projection
pursuit guided tour optimisation, there are various type of bases
involved: 1) The start basis; 2) The search bases that an optimiser
evaluated to produce the anchor bases; 3) The anchor bases that have the
highest index value in each iteration; 4) The interpolating bases on the
interpolation path; and finally 5) the end basis. The importance of
these basis differs and the most important ones are the start,
interpolating and end bases. Anchor and search bases can be turned on
with argument \texttt{details\ =\ TRUE}. Sometimes, two optimisers can
start with the same basis but finish with bases of opposite sign. This
happens because the projection is invariant to the sign difference of
the bases and so does the index value, however, this creates
difficulties for comparing the optimisers since the end bases will be
symmetric with respect to the origin. A sign flipping step is conducted
to flip the signs of all the bases in one routine if different
optimisations finish at opposite places.

Several annotations have been made to help understanding this plot. As
mentioned previously, the original basis space is a high dimensional
sphere and random bases on the sphere can be generated via the CRAN
package \CRANpkg{geozoo} \citep{geozoo}. Along with the bases recorded
during the optimisation and a zero basis, the parameters (centre and
radius) of the 2D space can be estimated. PCA is performed to get the
first two PC coordinates of all the bases and the centre of the circle
is the PC coordinates of the zero matrix, and radius is estimated as the
largest distance from the centre to the basis. The theoretical best
basis is known for simulated data and can be labelled to compare the how
close the end basis found by the optimisers is to the theoretical best.
Various aesthetics, i.e.~size, alpha and colour, are applicable to
emphasize critical elements and adjust for the presentation. For
example, anchor points and search points are less important and hence a
smaller size and alpha is used; Alpha can also be applied on the
interpolation paths to show the start to finish from transparent to
opaque.

Figure \ref{fig:toy-pca} shows the PCA plot of SA and PD for a 1D
projection problem. Both optimisers find the optimum but PD gets closer.
With PCA plot, one can visually appreciate the nature of these two
optimisers: PD first evaluates points in a small neighbourhood for a
promising direction, while SA evaluates points randomly in the search
space to search for the next target. There are dashed lines annotated
for SA and it describes the interruption of the interpolation, which
will be discussed in details in Section \ref{monotonic}.

@@

Animation is another display to show how the search progresses from
start to finish in the space and an \texttt{animate\ =\ TRUE} argument
is used to enable the animation in the PCA plot. Figure
\ref{fig:toy-pca-animated} shows six frames from the animation of the
PCA plot in Figure \ref{fig:toy-pca}. An additional piece of information
one can learn from this animation is that SA finds its end basis quicker
than PD since SA finishes its search in the 5th frame while PD is still
making more progression.

@@

As mentioned previously, the original \(p \times d\) dimension space can
be simulated via randomly generated bases in \CRANpkg{geozoo}
\citep{geozoo} package. While the PCA plot projects the bases from the
direction that maximises the variance, the tour plot display of the
original high dimensional space from various directions using animation.
Figure \ref{fig:toy-tour} shows frames from the tour plot of the same
two optimisations in its original 5D space.

@@

Random variables with different distributions have been simulated and
the distributional form of each variable is presented in Equations
\ref{eq:sim-norm} to \ref{eq:sim-x7}. Variable \texttt{x1}, \texttt{x8},
\texttt{x9} and \texttt{x10} are normal noise with zero mean and unit
variance and \texttt{x2} to \texttt{x7} are normal mixtures with varied
weights and locations. All the variables have been scaled to have an
overall unit variance before projection pursuit.

@@

Once the final basis has been found by an optimiser, one may want to
push further in the close neighbourhood to see if an even better basis
can be found. A polish search takes the final basis of an optimiser as
the start of a new guided tour to search for local breakthrough. The
polish algorithm is similar to the SA but with three distinctions: 1) a
hundred rather than one candidate bases are generated each time in the
inner loop; 2) the neighbourhood size is reduced in the inner loop,
rather than in the outer loop; and 3) three more termination conditions
have been added to ensure the new basis generated is distinguishable
from the current one in terms of the distance in the space, percentage
change in the index value, and neighbourhood size:

@@

The holes index function used for all the examples before this section
produces a smooth interpolation, while this is not the case for all the
indices. A 1D projection function, \(I^{nk}(n)\), compares the projected
data, \(\mathbf{Y}_{n \times 1}\), to a randomly generated normal
distribution, \(\mathcal{N}_{n \times 1}\), based on the Kolmogorov
test. Let \(F_{.}(n)\) be the ECDF function, with two possible
subscripts \(Y\) and \(\mathcal{N}\) representing the projected and
randomly generated data, and \(n\) denoting the number of observation,
the Normal Kolmogorov index, \(I^{nk}(n)\), is defined as:

\[I^{nk}(n) = \max \left[F_{Y}(n) - F_{\mathcal{N}}(n)\right]\] With a
non-smooth index function, two research questions are raised:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  whether any optimiser fails to optimise this non-smooth index; and
\item
  whether the optimisers can find the global optimum despite the
  presence of a local optimum
\end{enumerate}

@@

Figure \ref{fig:noisy-better-geo} presents the trace and PCA plots of
all three optimisers and as expected, none of the interpolated path is
smooth. There is barely any improvement made by PD, indicating its
failure in optimising non-smooth index. While SA and SAJO have managed
to get close to the index value of the theoretical best, trace plot
shows that it takes SAJO much longer to interpolate towards the final
basis. This long interpolation path is partially due to the fluctuation
in the early iterations, where the SAJO tends to generously accept
inferior bases before concentrating near the optimum. The PCA plot shows
the interpolation path and search points excluding the last termination
iteration. {[}A sentence for PD{]}. Comparing the amount of random
search done by SA and SAJO, it is surprising that SAJO doesn't carry
that many of sampling as SA. Combine the insight from both the trace and
PCA plot, one can learn the two different search strategies by SA and
SAJO: SA frequently samples in the space and only make a move when an
improvement is guaranteed to be made, while SAJO first broadly accepts
bases in the space and then starts the extensive sampling in a narrowed
subspace. The specific decision of which optimiser to use will depend on
the index curve in the basis space but if the basis space is non-smooth,
accepting inferior bases at first, as what SAJO has done, can lead to a
more efficient search, in terms of the overall number of points
evaluated.

@@

The next experiment compares the performance of SA and SAJO when a local
maximum exists. Two search neighbourhood sizes: 0.5 and 0.7 are compared
to understand how a large search neighbourhood would affect the final
basis and the length of the search. Figure \ref{fig:kol-result} shows 80
paths simulated using 20 seeds in the PCA plot, faceted by optimiser and
search size. With SA and a search size of 0.5, despite being the
simplest and fastest, the optimiser fails in three instances where the
final basis lands neither near the local nor the global optimum. With a
larger search size of 0.7, more seeds have found the global maximum.
Comparing SA and SAJO for a search size of 0.5, SAJO doesn't seem to
improve the final basis found, despite having longer interpolation
paths. However, the denser paths near the local maximum is an indicator
that SAJO is working hard to examine if there is any other optimum in
the basis space but the relative small search size has diminished its
ability to reach the global maximum. With a larger search size, almost
all the seeds (16 out of 20) have found the global maximum and some
final bases are much closer to the theoretical best, as compared to the
three other cases. This indicates that SAJO, with a reasonable large
search size, is able to overcome the local optimum and optimise close
towards the global optimum.

@@

The implementation of this projection has been divided into two
packages: the data collection object is implemented in the existing CRAN
package \CRANpkg{tourr} \citep{tourr} while the optimiser diagnostics
have been implemented in a new package, \pkg{ferrn}. When a guided tour
is run, the users can choose to collect the data by binding
\texttt{animate\_*()} to a variable. Once the data object has been
obtained, the \pkg{ferrn} package provides diagnostic plots shown in
Section \ref{vis-diag}. The main functions of the \pkg{ferrn} package
functionality is listed below.

@@

  \begin{itemize}
  \tightlist
  \item
    \code{explore\_trace\_search()} produces summary plots in Figure
    \ref{fig:toy-search}
  \item
    \code{explore\_trace\_interp()}produces trace plots for the
    interpolation points in Figure \ref{fig:toy-interp}
  \item
    \code{explore\_space\_pca()} produces the PCA plot of projection
    bases on the reduced space in Figure \ref{fig:toy-pca}. Animated
    version in Figure \ref{fig:toy-pca-animated} can be turned on via
    the argument \texttt{animate\ =\ TRUE}.
  \item
    \code{explore\_space\_tour()} produces animated tour view on the
    full space of the projection bases in Figure \ref{fig:toy-tour}.
  \end{itemize}
\item
  \code{get\_*()} extracts and manipulates certain components from the
  existing data object.

@@

  \begin{itemize}
  \tightlist
  \item
    \code{bind\_theoretical()} binds the theoretical best observation in
    simulated experiment
  \item
    \code{bind\_random()} binds randomly generated bases in the
    projection bases space to the data object
  \item
    \code{bind\_random\_matrix()} binds randomly generated bases and
    outputs in a matrix format
  \end{itemize}
\item
  \code{add\_*()} create ggprotos for different components for the PCA
  plot

  \begin{itemize}
  \tightlist
  \item
    \code{add\_anchor()} is a wrapper for plotting anchor bases
  \item
    \code{add\_anno()} is a wrapper for annotating the symmetry of start
    bases
  \item
    \code{add\_dir\_search()} is a wrapper for plotting the directional
    search bases with magnified distance
  \item
    \code{add\_end()} is a wrapper for plotting end bases
  \item
    \code{add\_interp()} is a wrapper for plotting the interpolation
    path
  \item
    \code{add\_interp\_last()} is a wrapper for plotting the last
    interpolation bases for comparing with target bases when
    interruption is carried
  \item
    \code{add\_interrupt()} is a wrapper for linking the last
    interpolation bases with target ones when interruption is carried
  \item
    \code{add\_search()} is a wrapper for plotting search bases
  \item
    \code{add\_space()} is a wrapper for plotting the circular space
  \item
    \code{add\_start()} is a wrapper for plotting start bases
  \item
    \code{add\_theo()} is a wrapper for plotting theoretical best bases,
    if applicable
  \end{itemize}
\item
  Utilities

  \begin{itemize}
  \tightlist
  \item
    \code{theme\_fern()} and \code{format\_label()} for better display
    of the grid lines and axis formatting
  \item
    \code{clean\_method()} for clean up the name of the optimisers
  \item
    \code{botanical\_palettes} is a collection of colour palettes from
    Australian native plants Quantitative palettes include daisy,
    banksia and cherry and sequential palettes contain fern and acacia
  \item
    \code{botanical\_pal()} as the colour interpolator
  \item
    \code{scale\_color\_*()} and \code{scale\_fill\_*()} for scaling the
    colour and fill of the plot
  \end{itemize}
\end{itemize}
